Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-10-01 paper.qmd
Updated Figure 10, nicely filtered out long edges outside the structure which showing in a small cluster, still need to work on how to remove long edges within the clutser (which is the most suitable benchmark)

@@

distance_two_curvy_dist <- distance_two_curvy_dist |>
  count(bin) |>
  mutate(prop = n/sum(n))

@@

distance_hist <- ggplot(
  distance_two_curvy_dist, 
  aes(x = as.numeric(bin),
      y = prop)) +
  geom_line() +
  scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1]),
                       expression(5*a[1]),
                       expression(6*a[1]),
                       expression(7*a[1]),
                       expression(8*a[1])
                     )) +
  geom_vline(xintercept = 2,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  geom_vline(xintercept = 4.5,
             linetype="solid",
             color = "#bdbdbd",
             linewidth=1) +
  ylab("Proportion of the number of edges") +
  xlab(expression(d^{(2)})) +
  #ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

>>>>

### Compute highD distance
dist_vec <- proxy::dist(x = df_bin_two_curvy2[, -1], method = "Euclidean") |> as.vector()

from_vec <- c()
to_vec <- c()
num_obs <- 1:(NROW(df_bin_two_curvy2) - 1)

for (obs in num_obs) {

  from_val <- rep(obs, (NROW(df_bin_two_curvy2) - obs))
  if ((obs + 1) <= NROW(df_bin_two_curvy2)) {
    to_val <- (obs + 1):NROW(df_bin_two_curvy2)
  }
  from_vec <- append(from_vec, from_val)
  to_vec <- append(to_vec, to_val)

dist_highd <- tibble::tibble(from = from_vec, to = to_vec, dist_highd = dist_vec)

@@

distance_two_curvy_dist <- left_join(distance_two_curvy_dist, dist_highd, by = c("from", "to"))

distance_hist <- ggplot(
  distance_two_curvy_dist, 
  aes(x = as.numeric(bin),
      y = dist_highd)) +
  geom_point(alpha = 0.5) +
  geom_jitter() +
  scale_x_continuous(breaks = sort(unique(as.numeric(distance_two_curvy_dist$bin))),
                     labels = c(
                       expression(a[1]),
                       expression(2*a[1]),
                       expression(3*a[1]),
                       expression(4*a[1]),
                       expression(5*a[1]),
                       expression(6*a[1]),
                       expression(7*a[1]),
                       expression(8*a[1])
                     )) +
  # geom_vline(xintercept = 2,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  # geom_vline(xintercept = 4.5,
  #            linetype="solid",
  #            color = "#bdbdbd",
  #            linewidth=1) +
  ylab(expression(d^{(7)})) +
  xlab(expression(d^{(2)})) +
  #ggtitle("(d)") +
  theme_minimal() +
  theme(aspect.ratio = 0.75,
        panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        panel.grid.major.x = element_blank(),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line())

# distance_two_curvy_dist <- distance_two_curvy_dist |>
#   count(bin) |>
#   mutate(prop = n/sum(n))

@@

# distance_hist <- ggplot(
#   distance_two_curvy_dist, 
#   aes(x = as.numeric(bin),
#       y = prop)) +
#   geom_line() +
#   scale_x_continuous(breaks = as.numeric(distance_two_curvy_dist$bin),
#                      labels = c(
#                        expression(a[1]),
#                        expression(2*a[1]),
#                        expression(3*a[1]),
#                        expression(4*a[1]),
#                        expression(5*a[1]),
#                        expression(6*a[1]),
#                        expression(7*a[1]),
#                        expression(8*a[1])
#                      )) +
#   geom_vline(xintercept = 2,
#              linetype="solid",
#              color = "#bdbdbd",
#              linewidth=1) +
#   geom_vline(xintercept = 4.5,
#              linetype="solid",
#              color = "#bdbdbd",
#              linewidth=1) +
#   ylab("Proportion of the number of edges") +
#   xlab(expression(d^{(2)})) +
#   #ggtitle("(d)") +
#   theme_minimal() +
#   theme(aspect.ratio = 0.75,
#         panel.border = element_rect(fill = 'transparent'),
#         plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
#         panel.grid.major.x = element_blank(),
#         axis.ticks.x = element_line(),
#         axis.ticks.y = element_line())
